The Astropy Project

### 2017-10-21 main.tex
add general section at end of article

\section{Introduction} \label{sec:intro}
% first draft by Moritz (hamogu)
All astronomical research makes use of software in some way and
astronomy as a field has long supported the development of specialized
software tools for different astronomical tasks. Some of those
software packages are (or at least were) supported by large
institutions, e.g.\ IRAF \citep[developed at NOAO,][]{IRAF}, MIDAS
\citep[developed at ESO,][]{MIDAS}, or ds9 \citep[developed at
% * <bsipocz@gmail.com> 2017-09-28T21:05:00.569Z:
%
% > or ds9
% probably starlink is a better example as being a large package of software similarly to IRAF and MIDAS rather than a single tool.
%
% ^.
% hamogu: Feel free ot swap in Starlink, if you think that's more appropriate. I picked ds9 becase (a) it's wider known (e.g. in the X-ray sounity) and (b) because it is an example of a widely used tool with institutional support that is NOT a large package. TOPCAT would have been another example if we want to expand that list.
  SAO,][]{ds9}. These packages tend to address a wide range of users
and they typically provide some level of documentation and user
support. Other packages are developed foremost by individual research
groups and used mostly in work by members of this group. Those
packages can address more specialized needs, but are not necessarily
available to the community and even if they are, documentation can in
some cases be insufficient for new users and the quality of the
software, i.e. the robustness and the level of testing that has been
performed to ensure correctness, is not always clear. On the other
hand, there are also many widely used, well documented, high quality
software packages which are developed by individuals or small groups.

The implementation of astronomical software, be it for a package meant
for wide distribution or scripts and programs for a specific research
project, can be eased by the use of a library that provides core
functionality that is common to many astronomical tasks. Reading and
writing FITS files is one example where it is time
consuming to implement the FITS standard correctly and
programming is much faster when using an existing library perform this
task. Another example of such a common task is dealing with different
astronomical coordinate systems. The \astropy project aims to provide a
library of core astronomical functionality in the \python programming
language.

\python\footnote{\url{https://www.python.org/}} is an increasingly popular
general-purpose scripting language that is available under a permissive open
source software licence free of charge for all major operating systems. Stable
and well-developed packages provide support for array arithmetic
\citep[\package{numpy},][]{numpy}, a wide variety of functions for scientific
computing \citep[\package{scipy},][]{numpy}, and publication-quality plotting
\citep[\package{matplotlib},][]{matplotlib}. Tens of thousands of other packages
are available which can help with tasks that are not astronomy specific but
might be performed in the course of astronomical research, e.g., presenting
results on websites or interfacing with databases.

@@

The first public release of the \astropy package is described in
\cite{astropy}. Since then, the \astropy package has been
used in hundreds of projects, and the scope of the package as grown
considerably. At the same time, the community of astronomers
contributing to the project has grown tremendously and an ecosystem
of packages supporting or affiliated with the \astropy core has
developed. In this paper, we describe the current status of the
\astropy community, the \astropy core package, and the \astropy\
ecosystem.

We start by describing the way the \astropy project works and is
organized in \sectionname~\ref{sec:concepts}.  The project develops a core
package called \astropypkg (\sectionname~\ref{sec:core}) and several
separate packages that help setting up the infrastructure for testing,
documentation etc.~(\sectionname~\ref{sec:infrastructure}). In
\sectionname~\ref{sec:learning} we give an overview of the wealth of
resources created to teach new users the use of the \astropy packages
and how to contribute to the community. We end with a short vision for
the future of \astropy in \sectionname~\ref{sec:future}.

@@

\subsection{Concept of affiliated packages}
% it may be described in previous subsection, if not the Brigitta can try to
% write it up
%
\par A major part of the \astropy~project is the concept of
``Affiliated Packages''. An affiliated package is an astronomy-related
\python~package that is not part of the \astropypkg~core package, but
has requested to be included as part of the \astropy project's
community. These packages support the goals and vision of \astropy~of
improving code reuse, interoperability, and embracing good coding
practices such as testing and thorough documentation. \astropypkg{}
version 2.0 supports \python~2.7 and \python~3.4+, affiliated packages
are strongly encouraged to support at least \python~3.x.
%
\par Affiliated packages contain functionality that is more specialized, or
have license incompatibility, or have enough external dependencies (e.g. GUI
libraries) that limit their integration possibility into the core and thus
they are more suitable to be separate packages. Starting as an affiliate 
package is also seen as a good strategy for functionality that is aimed for
the \astropypkg~core but is currently undergo rapid development and can 
benefit from a different release cycle while stabilizing its API. 
%
\par In general we hope that becoming an affiliated package is seen as 
a good way for new and existing packages to gain exposure.
%
\inlinecomment{BMS}{TODO: write about the process of becoming affiliated, and project packages into the future section below}
%
%\subsection{Mixin}

@@

\inlinecomment{BMS}{Maybe move the command line tools into a separate subsection to highlight them (even though we mostly only have fits related scripts)? They are supposed to be used more frequently by individual users.}

@@

dropping \python 2 support, growths of affiliated packages
Summary
learning astropy 
\acknowledgments

>>>>

\section{Introduction} \label{sec:intro}
% first draft by Moritz (hamogu)
All astronomical research makes use of software in some way and
astronomy as a field has long supported the development of specialized
software tools for different astronomical tasks. Some of those
software packages are (or at least were) supported by large
institutions, e.g.\ IRAF \citep[developed at NOAO,][]{IRAF}, MIDAS
\citep[developed at ESO,][]{MIDAS}, or ds9 \citep[developed at
% * <bsipocz@gmail.com> 2017-09-28T21:05:00.569Z:
%
% > or ds9
% probably starlink is a better example as being a large package of software similarly to IRAF and MIDAS rather than a single tool.
%
% ^.
% hamogu: Feel free ot swap in Starlink, if you think that's more appropriate. I picked ds9 becase (a) it's wider known (e.g. in the X-ray community) and (b) because it is an example of a widely used tool with institutional support that is NOT a large package. TOPCAT would have been another example if we want to expand that list.
  SAO,][]{ds9}. These packages tend to address a wide range of users
and typically provide some level of documentation and user
support is provided. Other packages are developed foremost by individual research
groups and used mostly in work by members of the group. Those
packages can address more specialized needs, but are not necessarily
available to the community and even if they are, documentation can in
some cases be insufficient for new users and the quality of the
software, i.e.\ the robustness and the level of testing that has been
performed to ensure correctness, is not always clear. On the other
hand, there are also many widely used, well documented, high quality
software packages which are developed by individuals or small groups.

The implementation of astronomical software, be it for a package meant
for wide distribution or scripts and programs for a specific research
project, can be eased by the use of a library that provides core
functionality that is common to many astronomical tasks. Reading and
writing FITS files is one example where it is time
consuming to implement the FITS standard correctly and
programming is much faster when using an existing library to perform this
task. Another example of such a common task is dealing with different
astronomical coordinate systems. The \astropy project aims to provide a
library of core astronomical functionality in the \python programming
language.

\python\footnote{\url{https://www.python.org/}} is an increasingly popular
general-purpose scripting language that is available under a permissive open
source software licence free of charge for all major operating systems. Stable
and well-developed packages provide support for array arithmetic
\citep[\package{numpy},][]{numpy}, a wide variety of functions for scientific
computing \citep[\package{scipy},][]{numpy}, and publication-quality plotting
\citep[\package{matplotlib},][]{matplotlib}. Tens of thousands of other packages
are available which can help with tasks that are not astronomy specific but
might be performed in the course of astronomical research, e.g.\ presenting
results on websites or interfacing with databases.

@@

The first public release of the \astropy package is described in
\cite{astropy}. Since then, the \astropy package has been
used in hundreds of projects, and the scope of the package has grown
considerably. At the same time, the community of astronomers
contributing to the project has grown tremendously and an ecosystem
of packages supporting or affiliated with the \astropy core has
developed. In this paper, we describe the current status of the
\astropy community, the \astropy core package, and the \astropy\
ecosystem.

We start by describing the way the \astropy project works and is
organized in \sectionname~\ref{sec:concepts}.  The project develops a core
package called \astropypkg (\sectionname~\ref{sec:core}) and several
separate packages that help setting up the infrastructure for testing,
documentation etc.\ (\sectionname~\ref{sec:infrastructure}). 
We end with a short vision for
the future of \astropy in particular and astronomical software in general
in \sectionname~\ref{sec:future}.

@@

\subsection{Concept of affiliated packages}
% it may be described in previous subsection, if not the Brigitta can try to
% write it up
%
\par A major part of the \astropy~project is the concept of
``Affiliated Packages''. An affiliated package is an astronomy-related
\python~package that is not part of the \astropypkg~core package, but
has requested to be included as part of the \astropy project's
community. These packages support the goals and vision of \astropy~of
improving code reuse, interoperability, and embracing good coding
practices such as testing and thorough documentation. \astropypkg{}
version 2.0 supports \python~2.7 and \python~3.4+, affiliated packages
are strongly encouraged to support at least \python~3.x.
%
\par Affiliated packages contain functionality that is more specialized, or
have license incompatibility, or have enough external dependencies (e.g. GUI
libraries) that limit their integration possibility into the core and thus
they are more suitable to be separate packages. Starting as an affiliate 
package is also seen as a good strategy for functionality that is aimed for
the \astropypkg~core but is initially undergoing rapid development and can 
benefit from a different release cycle while stabilizing its API. 
%
\par In general we hope that becoming an affiliated package is seen as 
a good way for new and existing packages to gain exposure.
%
\inlinecomment{BMS}{TODO: write about the process of becoming affiliated, and project packages into the future section below}
%
%\subsection{Mixin}

@@

\inlinecomment{BMS}{Maybe move the command line tools into a separate subsection to highlight them (even though we mostly only have fits related scripts)? They are supposed to be used more frequently by individual users.}
\inlinecomment{hamogu}{MY feeling is that this section is too detailed already. If I had written it, I would not have mentioned the scripts at all.}

@@

dropping \python 2 support, growths of affiliated packages
Summary
learning astropy

\section{Outlook}
\label{sec:outlook}
\inlinecomment{hamogu}{This may be merged into the section above once
  that is fleshed out a little more.}  As shown above the development
of the \astropypkg package is making good progress, helping
astronomers worldwide to perform many daily tasks such as planning
observations, analyzing data or simulation results, and writing
publications. The strong emphasis that the \astropy project puts on
reliability and code correctness helps users to trust the calculations
performed with \astropypkg and to publish reproducable results. More
functionality is added to the core package and more and more
affiliated packages support more specialized needs.

At the same time, \astropy is spreading awareness of best practices in
software development. This is important since most practicing
astronomers received their education without being taught in computer
science and software development although a major fraction of almost
every astronomer's workload today is related to software use and
development. The \astropypkg package leads by example, showing all interested
astronomers how modern tools like git verion control or continuous
integration (code is tested after every change to the source) can
increase the quality and accessability of astronomical software
without overly complicating the development cycle. All submitted code
is reviewed by at least one, but typically more, members of the
astropy community, providing feedback and allowing contributors to
improve their skill is developing software without the need for their
institutions to provide dedicated teaching recources. Thus, the
\astropy project does not only build the tools needed to perform
productive astronomical research in an era of increasingly large and
complex datasets, it also helps to prepare the current and the next
generation of researchers with the knowledge to adequatly use those
tools.

\acknowledgments
